### 2023-10-06 LaTeX_paper/paper.tex
finish review of section 6

\documentclass{article}

% Required for inserting images
\usepackage{graphicx}

% to have precise image placement
\usepackage{float}

% to insert hiperlinks
\usepackage[colorlinks=true,linkcolor=cyan]{hyperref}

% to have better references
\usepackage[noabbrev]{cleveref}

% to use prefix for references
\usepackage{etoolbox}

% to set the margin of the page
\usepackage[a4paper, total={6.5in, 9in}]{geometry}

% to make indentations
\usepackage{changepage}

% for sub figures
\usepackage{subcaption}

% to allow putting description text to te side of an image
\usepackage{wrapfig}

\usepackage{multirow}

% rule for reference to fugures
\pretocmd{\thefigure}{F}{}{}

% for the table
\usepackage[table,xcdraw]{xcolor}

% code highlighter
\usepackage{listings}

\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{orage}{rgb}{0,0.26,0.99}

\lstset{frame=tb,
  language=python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{orage},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

% to include svg images
%\usepackage{svg}

% settings for how to indent the page when needed
\newenvironment{indented_section}
  {\adjustwidth{3em}{0pt}}
  {\endadjustwidth}

\title{Use Of Jigsaw Puzzle Solving Algorithms In The Real World}
\author{Luca Sartore}
\date{May 2023}

\begin{document}

\maketitle

\newpage

% make the table of content without the hiperlink color
{
  \hypersetup{linkcolor=black}
  \tableofcontents
}

\newpage

\section{Abstract}
This paper examines the Jigsaw puzzle problem,
which has already been extensively studied in the literature.
Several algorithms with efficient time complexity have been developed
specifically for ``digital''~\ref{fig:figure_digital_puzzle} jigsaw puzzles. However, the counterpart:
the ``real world' jigsaw puzzle, has received relatively little attention
in the literature. While some small-scale projects have addressed this issue,
we were unable to find any papers discussing it.

The puzzle problem can be divided into two separate components: the ``Comparator''~\ref{document:comparator}
whose objective is to determine whether two sides of two pieces fit together,
and the ``Solver''~\ref{document:solver} whose objective is to use the information provided
by the comparator to find the best way to fit some pieces together.

Solving the problem in the real world has proven to be more challenging compared
to the digital realm.
This is because all measurements in the real world are noisier than those
in the digital world, making it challenging to obtain accurate data from the Comparator.
The Solver poses its own set of challenges, as it is an example of a
multiobjective optimization problem. Specifically, one objective is to have
the fastest algorithm possible,
while the other objective is to create an algorithm that performs well with
imprecise data. The latter objective is particularly important when solving
real-world puzzles.

In the first half of this paper, a Comparator is developed, and its accuracy
is evaluated.
The data are then used to demonstrate its incompatibility with the current
state-of-the-art algorithm. At this juncture,
two options remain to solve the problem:
either create a superior Comparator or design a
Solver that can operate with noisier data.

The most promising approach to improving the Comparator
involves utilizing machine learning, but this requires a substantial amount of data.
Conversely, creating a better Solver is a less straightforward task.
However, if successful, it could potentially serve as a labeling tool
to generate the dataset needed to train a superior Comparator.

The second part of this paper focuses on developing a
new Solver algorithm specifically designed to work with noisy data.
While the Solver created may not be considered state-of-the-art,
as its time complexity ultimately ends up being worse than the current
best algorithms, it could still play a fundamental role in solving the problem.
As mentioned earlier, it could be employed to solve many small-sized
puzzles and use the results to train a better Comparator.

In conclusion, this paper outlines the additional challenges that the real-world
puzzle problem presents compared to its digital counterpart.
Additionally, it lays the groundwork for a superior machine learning-based Comparator.

\section{Introduction}
\subsection{Classification}
This paper will focus on Type 2 puzzles. A Type 2 puzzle is one where the position and orientation of each piece are unknown.
\subsection{Digital vs Real-World Jigsaw Puzzles}

There is  another important distinction between different types of puzzles.
They can be divided into “digital”~\ref{fig:figure_digital_puzzle} and “real world”~\ref{fig:figure_real_puzzle} jigsaw puzzles.
The difference between the two is quite self explanatory, and can be observed in the respective pictures.
\label{document:DigitalVSReal}

% figure of digital jigsaw puzzle    
\begin{figure}[H]
    \caption{An example of a “digital” jigsaw  puzzle}\label{fig:figure_digital_puzzle}
    \centering
    \includegraphics[height=0.25\textwidth]{pictures/digital_puzzle.png}
\end{figure}

% figure of real jigsaw puzzle    
\begin{figure}[H]
    \caption{An example of a ``real world'' jigsaw  puzzle}~\label{fig:figure_real_puzzle}
    \includegraphics[height=0.25\textwidth]{pictures/real_puzzle.jpg}
    \centering

\end{figure}

The reason this distinction is important is because,
despite the generic concept of the puzzle not changing,
obtaining accurate matches of a piece's characteristics
is far easier with a digital puzzle,
since there are far fewer things that can go wrong.
For example, a piece's shape may be distorted by
a low-quality cut, as seen in picture \ref{fig:figure_measurement_error}

% figure of a measurement error
\begin{figure}[H]
    \caption{An example of what can go wrong when dealing with the real world}\label{fig:figure_measurement_error}
    \includegraphics[height=0.25\textwidth]{pictures/example_bad_piece.jpg}
    \centering
\end{figure}

\section{Previous Literature}
This section will analyze three different algorithms that have been
proposed as solutions for type 2 puzzles.
The objective is to understand the strengths and weaknesses of each one,
in order to build up knowledge that will be useful for the next sections.

\subsection{General Structure Of The Algorithms}

All the algorithms that will be analyzed are composed of 3 sub algorithms:

\begin{indented_section}

    \subsubsection{The Splitter:} This component takes as input one or more images
    containing all the pieces. It then separates all the pieces from each other,
    and split each piece into his four sides.\label{document:splitter}

    \subsubsection{The Comparator:} This component compares each side with all the
    others, in order to understand whether they match or not.\newline
    There are two distinct kinds of “Comparator” algorithms:
    The “Binary Comparison” and the “Non Binary Comparison”.
    As the name suggests when comparing two sides with a “Binary Comparison”
    the result can either be 0 (they do not match) or 1 (they match).
    In contrast a “Non Binary Comparison” can give any value between 0 and 1.
    This allows states of uncertainty to be represented.\label{document:comparator}

    \subsubsection{The Solver:} This component utilizes the information provided by the
    Comparator~\ref{document:comparator}, and attempts to find a solution
    (i.e.\ a position and an orientation for each piece)
    that is the most likely to be correct.\label{document:solver}

\end{indented_section}

\subsection{Solving Jigsaw Puzzles By The Graph Connection Laplacian~\cite{GCL}}
This algorithm falls within the ``binary comparison'' category~\ref{document:comparator}.
And can be used to understand the strength and weaknesses
of this approach.\newline
The primary advantage of this approach lies in its speed;
the utilization of binary comparison permits specific optimizations,
particularly leveraging graph theory techniques.\newline
The algorithm exhibits a time complexity of approximately \(O(N^2)\),
a value consistent with the theoretical minimum for
jigsaw puzzle solving, as outlined in the study
``No easy puzzles: Hardness results for jigsaw puzzles~\cite{ON2Claim}''.\newline
However, a potential drawback may arise in terms of accuracy.
This stems from the fact that, with only two states
(matching or not matching),
certain details may be lost when compared to a non-binary comparison. 
To quantify accuracy the paper used the ``neighbors compariso'' metric,
which ``calculates the percentage of pairs of image patches that are matched correctly''.\label{document:GCL}

\subsection{A Genetic Algorithm-Based Solver for Very Large Jigsaw Puzzles~\cite{GA}}
This paper employs genetic algorithms to address the
jigsaw puzzle problem, yielding promising results.\newline
The algorithm seems to have a \(O(N^2)\) time complexity,
with execution time slightly lower than the previous example
(However, it remains uncertain which one is faster, as the papers did not specify
the particular hardware configuration used).\newline
The paper adopts the same evaluation method for accuracy as the previous example,
namely neighbor comparison, facilitating direct comparison between the two approaches.
This algorithm falls under the category of “Non binary comparison”~\ref{document:comparator}
a distinction that theoretically grants it an advantage due to the increased granularity
of input data.
Unfortunately this advantage does not compensate for the worst precision
of the algorithm itself, and the accuracy results are equivalent,
if not slightly worse than the previous algorithm.\newline
It is important to keep in mind that for the very nature of genetic algorithms,
it might be possible that the accuracy would have been better if they had allowed
the algorithm to run for more generations.\label{document:GA}

\subsection{Computer Vision Powers Automatic Jigsaw Puzzle Solver~\cite{Abto}}
Solutions~\ref{document:GCL} and~\ref{document:GA} can be considered ``state-of-the-art''
for digital jigsaw puzzles.
However, as already introduced in section ~\ref{document:DigitalVSReal},
translating these solutions to the real world can be challenging.\newline
This article represents one of the most comprehensive and well-documented
implementations for real-world puzzles. However,
it's important to note that referring to this approach as a 'solution'
may be misleading, as it primarily identifies the top eight matching
pieces for a specific point and subsequently relies on user input to make
the final placement decision
This aspect renders the algorithm non-autonomous..\newline
Even disregarding this significant detail,
conducting an analysis of this solution has proven to be challenging,
as the article leans more towards a publicity stunt than an academic paper.
In fact, the code is not open source, and information regarding the time needed
to solve the puzzle is absent. The article only showcases a single puzzle solved,
with dimensions of 9\(\times\)6.\newline
Despite these limitations, this article still stands as one of the best examples of a solution
for real-world puzzles.

\section{The Objective Of This Paper}\label{document:objective}
After a review of the existing literature on this problem, some conclusions could be drawn:
The "digital" jigsaw puzzle dilemma has been extensively addressed and is already
well-optimized. Conversely, the "real-world" counterpart has largely been
overlooked within the academic community.\newline
The first objective for this paper is to fill this void, and been the first formal paper
to analyze the real-world jigsaw puzzle problem.\newline
The second objective is to use the forthcoming research to create a labeling tool.\newline
This objective is motivated by the fact that the biggest challenge for real world puzzles is creating a
reliable comparator~\ref{document:comparator}, that could provide accurate information even with noisy 
input images, that are obtained from real-world measure.\newline
It is easy to see how a machine learning based comparator would probably be the best solution, 
but this poses problems, in fact having to manually scan and label thousands of individual puzzle
pieces to train the model would take ages.
Instead an automatic tool would allow to scan a small $4\times4$ puzzle in one shot,
solve it, and return the labeled data in just minutes.

\section{The Setup}

To fulfill the objectives of this paper,
it is essential to have the necessary data for analysis.
This section will encompass the input data utilized,
as well as the Splitter~\ref{document:splitter}
and Comparator~\ref{document:comparator} algorithms.

\subsection{The Puzzle To Solve}

The first step involves selecting a real-world
puzzle for solving. It has been decided to utilize a
classic 1000-piece puzzle measuring 66cm x 50cm.
A printer's scanner is employed to digitize the individual pieces.\newline

The puzzle has been scanned in various sub-sizes to facilitate the testing of different configurations.
A resolution of 1200 pixels per inch (ppi) has been employed to ensure a sharp and precise
depiction of the pieces, as seen in image~\ref{fig:figure_measurement_error}.
The puzzle has been scanned backward, using a black background.

This choice has been dictated by the fact that the scanner used has the flashlight
slightly off-centered from the scanner's sensor.
And this causes a shadows to be cast on one side of the pieces.
A black background has been chosen to solve the problem of the shadow.
And then the pieces has been flipped backward, since many of them where black,
and it was difficult to separate them from the background.
In the future it would be possible to scan the pieces in the correct orientation,
by using a different scanner.

\subsection{Splitting The Pieces}

As introduced in section~\ref{document:splitter},  the initial stage in puzzle solving involves
splitting an image composed of many pieces into individual ones.
This is relatively straightforward, and can be done with the following steps:

\begin{itemize}
  \item Apply a threshold and transform an RGB image into a binary mask.
  \item Find all individual Blobs
  (a Blob is defined as a set of connected pixels with the same color) and for each one:

  \begin{itemize}
  \item Calculate the area (the number of pixels the Blob is made of)

  \item Add it to the set of pieces only if his area in within a certain range (to avoid small pieces of dust ending up counted as pieces)

  \end{itemize}

\end{itemize}

\subsection{Splitting The Sides}

The next step following the general structure~\ref{document:splitter}
is to split a piece into
his 4 sides, this could be done with the following steps:

\begin{itemize}

  \item \textbf{Remove the holes from the piece:}\newline
  To remove the holes from the piece, the program employs a three-step process. Initially it
  computes the convex hull of the piece~\ref{fig:s_s_ch}.
  Subsequently it calculates the difference between the convex hull and the original image,
  resulting in an ``image with filler areas''~\ref{fig:fig:s_s_og_minus_ch}.
  It can be noted that these filler areas have distinct shapes,
  depending on what has generated them (either a ``hole'' or a ``knob'').
  Therefore they can be categorized.
  Finally the filler areas are added back to the only if
  they originated from a hole. This results in an image where the holes
  have been filled~\ref{fig:s_s_no_holes}

  \item \textbf{Remove the knobs from he piece:}\newline
  In order to remove the knobs from the piece, the image without holes
  undergoes an erosion~\ref{fig:s_s_erosion} followed by a dilation~\ref{fig:s_s_dilatation}.
  This process results in an image that resembles the original, but without any knobs.
  Subsequently, the program identifies the pixels that are white in the image without holes~\ref{fig:s_s_no_holes},
  but not in the eroded and dilated version~\ref{fig:s_s_dilatation}.
  This leads to the formation of a set of blobs ~\ref{fig:s_s_no_holes_minus_expansion}.
  It is important to note that the shape of a blob can vary significantly depending on whether it originated
  from an ``angle'' or a ``knob'' and thus they can be classified accordingly.
  Following this, the program proceeds to eliminate from the hole-free
  image all pixels that are in close proximity to a blob generated by a knob.
  This operation yields an image that is both devoid of knobs and holes~\ref{fig:s_s_no_knobs}.

  \item \textbf{Find the coordinates of the four corners:}\newline
  In order to determine the coordinates of the corners,
  the program employs a three-step approximation method.
  Initially, it identifies the minimum enclosing rectangle and
  utilizes the coordinates of its corners
  as the first approximation~\ref{fig:s_s_min_enc_rec}.\newline
  For the second approximation, the program selects,
  from the white pixels in close proximity to the first approximation,
  the pixel that contains the highest number of black pixels
  within a specified range.
  Then it considers only the pixels that are within a certain
  distance of the second approximation, and finds the minimum
  enclosing triangle of sed pixels~\ref{fig:s_s_min_enc_triangle}.\newline
  The corner of the triangle that is nearest to the second
  approximation serves as the third and final approximation
  for the corner of the puzzle piece.
  The position of the four corners can be seen in image~\ref{fig:s_s_corners_found}.

  \item \textbf{Split each piece into four sides using the previously found coordinates:}\newline
	After the program has successfully identified the four corner coordinates,
  it can straightforwardly derive each of the four sides.
  An example of a single side can be seen in image F4l.~\ref{fig:s_s_a_piece_side}.
\end{itemize}

\begin{figure}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/original_piece.jpeg}
    \caption{Original binary mask of a piece.}
    \label{fig:s_s_og}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/remove_holes_convex_hull.png}
    \caption{Convex Hull of a piece.}
    \label{fig:s_s_ch}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/remove_holes_filler_areas.png}
    \caption{Difference:~\ref{fig:s_s_ch} \textminus~\ref{fig:s_s_og}.}
    \label{fig:fig:s_s_og_minus_ch}
  \end{subfigure}
  \vspace{1cm}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/piece_with_no_hole.png}
    \caption{Piece~\ref{fig:s_s_og} with holes removed.}
    \label{fig:s_s_no_holes}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/remove_knobs_erosion.png}
    \caption{Mask~\ref{fig:s_s_no_holes} eroded.}
    \label{fig:s_s_erosion}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/remove_knobs_expansion.png}
    \caption{Mask~\ref{fig:s_s_erosion} dilated.}
    \label{fig:s_s_dilatation}
  \end{subfigure}
  \vspace{1cm}
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/remove_knobs_knob_pixels.png}
    \caption{Difference:~\ref{fig:s_s_no_holes} \textminus~\ref{fig:s_s_erosion}.}
    \label{fig:s_s_no_holes_minus_expansion}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/piece_with_no_knobs.png}
    \caption{Piece~\ref{fig:s_s_no_holes} with knobs removed.}
    \label{fig:s_s_no_knobs}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/find_corners_min_enclosing_rectangle.png}
    \caption{Min enclosing rectangle of~\ref{fig:s_s_no_knobs}.}
    \label{fig:s_s_min_enc_rec}
  \end{subfigure}
\end{figure}
\clearpage
\begin{figure}\ContinuedFloat
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/find_corners_min_enclosing_triangle.png}
    \caption{Min enclosing triangle of the bottom right corner of~\ref{fig:s_s_no_knobs}.}
    \label{fig:s_s_min_enc_triangle}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/corners_found.png}
    \caption{The corners of the piece~\ref{fig:s_s_og}.}
    \label{fig:s_s_corners_found}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.3\textwidth}
    \centering
    \includegraphics[width=\linewidth]{pictures/piece_side.png}
    \caption{One of the sides of the piece~\ref{fig:s_s_og}.}
    \label{fig:s_s_a_piece_side}
  \end{subfigure}

  \caption{Steps needed to split an image into his four sides.}
  \label{fig:splitting_sides}

\end{figure}

\subsection{Comparing The Sides}\label{document:my_comparator}
The last step in the setup, is to compare the sides that have been
generated before.\newline
In order to generate a compatibility shore,
the program compute the following steps:

\begin{itemize}
  \item For each side, it considers only the border, and  makes it thicker.
  \item It puts two sides attached to each other using the coordinates of the corners (In the same way a human would do to test if two pieces fit together).
  \item It calculates the “Or Area”; that is, the number of pixels that belong to the border of one side or the other.
  \item It calculates the “And Area”; that is, the number of pixels that belong to the border of both sides.
  \item It calculates the final shore, which is defined as: \(Shore = \frac{And \space Area}{Or \space Area}\), and goes from 0 to 1.
\end{itemize}

\begin{figure}[htbp]
  \centering
  \begin{minipage}[t]{0.44\textwidth}
    \vspace{3pt} % Ensure alignment of the top of the minipage
    \includegraphics[width=\textwidth]{pictures/side_comparation.png}
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.54\textwidth}
    \caption{\newline
    In this image we can see two sides that have been
    compared, one side is drowned in red, the other one in blue.
    In this picture, the “Or Area” is represented by all non
    black pixels; The “And Area” is represented by all the
    purple pixels.}
  \end{minipage}
\end{figure}
\clearpage

\section{Applying Already Existing Algorithms To The Real World}
With the initial setup completed, it is now possible to
test the existing algorithms with real-world puzzle. For this test,
the fist algorithm~\ref{document:GCL} is selected due to the comprehensive
details provided in the paper. Notably, it contains information
regarding the algorithm's accuracy in relation to the corruption rate,
which is crucial for this analysis.

The conducted tests reveal that the Comparator developed in this paper~\ref{document:my_comparator}
and the "Graph Connection Laplacian" solution ~\ref{document:my_comparator} are not compatible with each other.
A formal proof supporting this assertion is presented subsequently.

\subsection{Proof}

The comparison algorithm deceived in section~\ref{document:my_comparator} outputs a compatibility
shore that goes from 0 to 1.\newline
But the algorithm~\ref{document:GCL} takes as input a binary value.
Which means that it is necessary to apply a threshold to the input
data to convert them from float to bool.
\subsubsection{Input Data}
To find the optimal threshold a small 16 piece puzzle has been solved.
And the following 4 values have been measured, in function of the threshold:

\begin{itemize}
  \item \textbf{match\%:}\newline
  The probability that 2 randomly selected sides matches.

  \item \textbf{correct combination match\% (or c.c match\%):}\newline
  The probability that 2 randomly selected sides matches given that in the correct puzzle
  solution they are actually close to each other. Ideally this should be 1.

  \item \textbf{limit match\%:}\newline
  Given that this measure has been taken on a 16 piece puzzle,
  the match\% is highly influenced by the correct piece matching together.
  With a bigger puzzle the match\% would be lower (Given that the amount
  of correct couple of sides scales with N, while the amount of
  possible couple of sides scales with \(N^2\). Where N is the number of pieces).\newline
  Limit match\% is an approximation of what the match\%
  would be in the limit for \(N \rightarrow + \infty\).
  This value is obtained by subtracting to match\%
  the probability of choosing randomly the correct side,
  multiplied by the probability that the comparison returns true
  (which is c.c match\%). \newline
  Ideally limit match\% should be almost zero.
  \[limit \: math\% = math\% - \frac{c.c. \: mathc\%}{16 \times 4}\]

  \item \textbf{corruption rate:}\newline
  Is defined as: \(1-\sqrt{c.c \: match\%}\). Ideally this should be 0.

\end{itemize}

% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[H]
  \centering
  \begin{tabular}{
  >{\columncolor[HTML]{FFE599}}c 
  >{\columncolor[HTML]{B6D7A8}}c 
  >{\columncolor[HTML]{9FC5E8}}c 
  >{\columncolor[HTML]{F9CB9C}}c 
  >{\columncolor[HTML]{D9D2E9}}c }
  \cline{1-1}
  \multicolumn{1}{|c|}{\cellcolor[HTML]{F1C232}threshold} &
    \cellcolor[HTML]{6AA84F}match\% &
    \cellcolor[HTML]{3C78D8}c.c match\% &
    \cellcolor[HTML]{E69138}limit match\% &
    \cellcolor[HTML]{8E7CC3}corruption rate \\ \cline{1-1}
  0.000 & 1.000 & 1.000 & 0.984 & 0.000 \\
  0.025 & 0.408 & 1.000 & 0.392 & 0.000 \\
  0.050 & 0.408 & 1.000 & 0.392 & 0.000 \\
  0.075 & 0.408 & 1.000 & 0.392 & 0.000 \\
  0.100 & 0.408 & 1.000 & 0.392 & 0.000 \\
  0.125 & 0.406 & 1.000 & 0.391 & 0.000 \\
  0.150 & 0.400 & 1.000 & 0.385 & 0.000 \\
  0.175 & 0.386 & 1.000 & 0.371 & 0.000 \\
  0.200 & 0.364 & 1.000 & 0.348 & 0.000 \\
  0.225 & 0.335 & 1.000 & 0.319 & 0.000 \\
  0.250 & 0.298 & 1.000 & 0.282 & 0.000 \\
  0.275 & 0.253 & 1.000 & 0.237 & 0.000 \\
  0.300 & 0.201 & 1.000 & 0.185 & 0.000 \\
  0.325 & 0.158 & 1.000 & 0.142 & 0.000 \\
  0.350 & 0.125 & 0.944 & 0.110 & 0.028 \\
  0.375 & 0.100 & 0.944 & 0.085 & 0.028 \\
  0.400 & 0.080 & 0.944 & 0.065 & 0.028 \\
  \cellcolor[HTML]{FF0000}0.425 &
    \cellcolor[HTML]{FF0000}0.054 &
    \cellcolor[HTML]{FF0000}0.889 &
    \cellcolor[HTML]{FF0000}0.040 &
    \cellcolor[HTML]{FF0000}0.057 \\
  0.450 & 0.043 & 0.889 & 0.029 & 0.057 \\
  0.475 & 0.034 & 0.778 & 0.022 & 0.118 \\
  0.500 & 0.025 & 0.778 & 0.013 & 0.118 \\
  0.525 & 0.023 & 0.778 & 0.010 & 0.118 \\
  0.550 & 0.020 & 0.667 & 0.010 & 0.184 \\
  0.575 & 0.019 & 0.667 & 0.009 & 0.184 \\
  0.600 & 0.016 & 0.556 & 0.007 & 0.255 \\
  0.625 & 0.014 & 0.556 & 0.005 & 0.255 \\
  0.650 & 0.014 & 0.556 & 0.005 & 0.255 \\
  0.675 & 0.006 & 0.222 & 0.003 & 0.529 \\
  0.700 & 0.003 & 0.111 & 0.001 & 0.667 \\
  0.725 & 0.001 & 0.000 & 0.001 & 1.000 \\
  0.750 & 0.001 & 0.000 & 0.001 & 1.000 \\
  0.775 & 0.000 & 0.000 & 0.000 & 1.000 \\
  0.800 & 0.000 & 0.000 & 0.000 & 1.000 \\
  0.825 & 0.000 & 0.000 & 0.000 & 1.000 \\
  0.850 & 0.000 & 0.000 & 0.000 & 1.000 \\
  0.875 & 0.000 & 0.000 & 0.000 & 1.000 \\
  0.900 & 0.000 & 0.000 & 0.000 & 1.000 \\
  0.925 & 0.000 & 0.000 & 0.000 & 1.000 \\
  0.950 & 0.000 & 0.000 & 0.000 & 1.000 \\
  0.975 & 0.000 & 0.000 & 0.000 & 1.000 \\
  1.000 & 0.000 & 0.000 & 0.000 & 1.000
  \end{tabular}
  \end{table}

\begin{figure}[H]
    \caption{match metrics vs threshold used}\label{fig:match_metrics_vs_threshold_used}
    \centering
    \includegraphics[height=0.5\textwidth]{pictures/match_metrics_vs_threshold_used.png}
\end{figure}

Now, let's consider Algorithm \ref{document:GCL} tasked with solving a 500-piece puzzle
using our Comparator \ref{document:my_comparator}.
A threshold of 0.425 is chosen;
This implies a imit match\% of 0.04 and a corruption rate of 0.057.
What would this mean in terms of accuracy of the results?

\subsubsection{Definitions}\label{document:proof_definitions}
This set of definitions are useful to better understand the next section:

\begin{itemize}
  \item \textbf{Connection:}\newline
  When two sides of two different pieces are joined they form a connection.
  A solved puzzle with \(N\) pieces, (assuming it is squared) has around \(2N\) total connections.

  \item \textbf{Possible Combinations:}\newline
  A puzzle with N pieces can be put together in many different ways,
  (Even if not all of 	them actually fit together).
  The total number of possible combinations is \(N! \times 4^N\) where
  \(N!\) represent all possible way to combinate the pieces in a particular position, and
  \(4^N\) represents all possible orientations that each piece can have.
  \item \textbf{Possible Correct Combination:}\newline
  Is one of the possible combinations where all the connections are considered a match
  by the comparing algorithm.
  \item \textbf{Actual Correct Combination:}\newline
  An Actual Correct Combination is a Possible Combination that
  is also the correct solution of the original puzzle	
\end{itemize}

\subsubsection{Corruption Rate Analysis}
\begin{figure}[h]
  \caption{Neighbors comparison vs Corruption rate of algorithm \ref{document:GCL}. Image taken from pater \cite{GCL} page 34.}\label{fig:GCL_Corruption_Rate}
  \centering
  \includegraphics[height=0.35\textwidth]{pictures/corruption_rate.png}
\end{figure}

In Figure \ref{fig:GCL_Corruption_Rate}, it can be observed that at a corruption rate of 0.057,
a neighbors comparison metric of roughly 80\% can be anticipated.
This implies that, on average, one out of every five connections will be incorrect.
In the context of a puzzle comprising 500 pieces (equating to approximately 1000 connections),
an estimated 200 connections will be inaccurate on average.
It is reasonable to assert that this outcome is suboptimal.

\subsubsection{Limit Match\% Analysis}
Let's remember that \(limit\_match\% = 0.04\).

Given that in a puzzle with N pieces there are around \(2N \) connections,
the probability of one random combination to be a
possible correct combination is: \(0.04^{2N}\).

The total number of possible combinations are \(N! \times 4^N\),
which means that on average, in a puzzle with \(N\) pieces,
there will be \(0.04^{2N} \times N! \times 4^N \)
possible correct combinations.

For \( N = 500 \) the expected number of possible combinations is around \( 1.5 \times 10^{37} \).
Which makes the chance of finding the actual correct combination basically zero.
\subsubsection{Conclusion}
In this case both the match\% and the corruption rate are bad enough to make a correct reconstruction impossible.
To lower limit match\% is necessary to increase the threshold, but this would increase the corruption rate.
To lower the corruption rate is necessary to decrease the threshold, but this would increase the limit match\%.
This means that, regardless of the threshold chosen, our comparison algorithm \ref{document:my_comparator}
can not work with a state of the art solver \ref{document:GCL}.

\subsection{Solutions For The Problem}

There are two viable approaches to overcome this challenge.
It is either necessary to enhance the Comparator while continuing
to utilize the state-of-the-art solver,
or it is possible to re-design the solver algorithm and create one
specifically designed to operate with noisier data.

\subsubsection{Improve The Comparator}

Improving the comparator can shortly be done.
Some minor improvements can be made by spending some time
tweaking the current algorithm.
However, it is evident that for a task of this nature,
the most viable solution would likely involve
leveraging machine learning techniques.

\subsubsection{Improving The Solver Algorithm}

The puzzle dilemma is a ``multiobjective optimization problem'',
where one objective is optimizing the time complexity,
and the other is optimizing the accuracy of the algorithm
in the presence of corrupted data.
All the major algorithms proposed up to this point have prioritize speed.
However, it is conceivable, and even probable,
that an algorithm prioritizing accuracy over speed could prove
effective with our comparator

\subsubsection{The Path Forward}

Improving the comparator is likely the most effective long-term solution,
but it presents a significant challenge: manually testing if certain
pieces fit together, followed by scanning and labeling,
is a time-intensive process that could require hundreds,
if not thousands, of hours.

On the other hand, improving the solver may not be the optimal long-term
solution, as it is likely to result in a worse time complexity.
However, it could yield significant short-term benefits.

Notably, this approach does not necessitate the creation of an extensive dataset
and has the potential to facilitate its development.
In fact, it could have the capability to solve numerous small
puzzles and subsequently auto-label the connections
based on the solutions it generates.

Given the two options, it has been decided to focus on improving
the solver algorithm. This choice was driven by the belief
that creating a new algorithm, and subsequently using it to
generate a dataset, would still be a more time-efficient
approach compared to manually constructing a dataset.

\section{The Solver}\label{document:my_solver}
The Solver that has been created is not the main focus of this paper,
and cannot be considered state-of-the-art if compared with the current best
digital puzzle solver. But can probably be considered so if compared with current
``real world'' puzzle Solvers.

\subsection{Overview}
The algorithm itself is relatively simple.
It is a recursive procedure that begins by identifying all 2x2 sets of pieces that match by combining four individual pieces.
It then proceeds to identify all 4x4 sets of matching pieces by combining four 2x2 pieces, and so on.\newline

\subsection{Basic Components}
To solve the puzzle two fundamental object are used
\begin{itemize}
  \item SinglePiece: which contains information about a single puzzle piece.
  \item PieceGroup\(<\)T\(>\) where T can be either a PieceGroup or a SinglePiece\newline
  This element encompasses four elements of type T, each with its own position (top right, top left, bottom left, bottom right).
\end{itemize}
Both SinglePiece and PieceGroup have an ``orientation'' attribute that goes from 0 to 3~\ref{fig:orientation}.
\begin{figure}[H]
  \caption{orientation attribute of an object}\label{fig:orientation}
  \centering
  \includegraphics[height=0.3\textwidth]{pictures/orientation.png}
\end{figure}

\subsubsection{The Levels}

These objects are assigned levels:
\begin{itemize}
\item SinglePiece is at level 0, representing a 1x1 ``solution'' to a puzzle.
\item PieceGroup\(<\)SinglePiece\(>\) is at level 1, representing a 2x2 solution to a puzzle.
\item PieceGroup\(<\)PieceGroup\(<\)SinglePiece\(>>\) is at level 2, representing a 4x4 solution to a puzzle, and so forth.
\end{itemize}

\subsubsection{The Behaviors}

These basic objects also have the following shared behaviors:
\begin{itemize}
  \item 4 objects of level N can be combined to build an object of level N+1.
  \item One object of level N has 4 sides, each of which can be compared with a side of an element of level N. This results in a compatibility shore.
  \begin{itemize}
    \item The comparing shore for level 0 objects is just the shore given by the comparator~\ref{document:comparator}.
    \item The comparing shore for level N+1 objects is the average between two comparisons made with objects of level N
  \end{itemize}
\end{itemize}

\subsubsection{Combining four objects}\label{document:combining_four_objects}

When combining four objects into one, some comparisons are executed to ensure that the resulting piece is a valid  combination.
\begin{itemize}
  \item The right side of the object in the top left position is compared with the left side of the object in the top right position~\ref{fig:merge_pieces}(1).
  \item The bottom side of the objection the top right position is compared with the top side of the object in the bottom right position~\ref{fig:merge_pieces}(2).
  \item The left side of the objection the bottom right position is compared with the right side of the object in the bottom left position~\ref{fig:merge_pieces}(3).
  \item The top side of the objection the bottom left position is compared with the bottom side of the object in the top left position~\ref{fig:merge_pieces}(4).
\end{itemize}

\begin{figure}[H]
  \caption{Example of a combination of four objects in to one}\label{fig:merge_pieces}
  \centering
  \includegraphics[height=0.3\textwidth]{pictures/merge_pieces.png}
\end{figure}

In order for the created object to be considered valid all of the comparisons must have a shore higher
than the constant MIN\_SHORE\_SINGLE\_SIDE. And the average of the four comparisons must be greater
than the constant MIN\_SHORE\_PIECE\_GROUP.

The process also checks that all individual pieces inside a newly
created object are different (because obviously is impossible to reuse the same piece twice to solve a puzzle)
\subsection{The Algorithm}

As mentioned earlier, the algorithm is implemented with a recursive function.
The function is generic, and can take as input objects of any level.

A Function of level N takes as input a list of objects of level N,
and creates all possible objects with level N+1.
Then it uses the newly created objects to generate a recursive call on a version
of itself with level N+1.
This process is repeated until the desired size is reached.

\begin{minipage}{\textwidth}
\begin{lstlisting}
  # This function returns a list of object that match with `piece_to_match'
  # along the direction `direction'.
  # It also filters the data to return only pieces with an index greater than `index_filter'
  # This function is executed in constant time, since it relies on some pre-processing of the data
  def get_matches(piece_to_match, direction, index_filter):
      ...

  def solve<T>(pieces: List<T>):

      number_of_pieces = pieces.length()

      found_pieces = List::<PieceGroup<T>>::new()

      # this function compare all pieces with the others, and build some data structures
      # that are used from the function `get_matches'
	   # the time complexity is o(N^2)
      do_pre_processing(pieces)

      # test all possible combinations for the top left piece
      for top_left_index in 0..number_of_pieces:
	       for top_left_orientation in 0..4:

          # get the top left piece
          top_left = get_piece(top_left_index, top_left_orientation)

          # test all possible combinations of top_right, bottom_right and bottom_left
          for top_right in get_matches(top_left, RIGHT, top_left_index):
	           for bottom_right in get_matches(top_right, DOWN , top_left_index):
                  for bottom_left in get_matches(bottom_right, LEFT, top_left_index):

                      # try to create the new piece group
                      new_element = PieceGroup::new(top_left, top_right, bottom_right,bottom_left)

                      if new_element.is_valid():
                          # if the 4 pieces can fit together the newly created piece
                          # is added to the list of found pieces
                          found_pieces.add_element(new_element)
                      else:
                          # otherwise the program check what the case of the 
                          # invalid creation is, and generate a continue
                          # on the respective loop, to speed up the operations
                          continue_failure_cause_loop()

      # at the end a recursive call is generated 
      return solve::<PieceGroup<T>>(found_pieces)

\end{lstlisting}
\end{minipage}
The procedure begins by testing all possible pieces and orientations as the top\_right piece.
Then continues by finding all the other 3 pieces in an efficient way thanks to get\_matches.
When the program has an hypothesis on the 4 pieces that can build a new piece it tries to create it.\newline
The creation process does some checks, and if everything is successful it adds the piece to the list of found pieces.
If the creation is not successful it figures out which piece has caused the failure, and moves to the next iteration of the respective loop.\newline
When all combinations have been found, the function makes a recursive call and the process is repeated.\newline
The process ensures that the piece in the top\_left is always the one with the lower index.
This is done to avoid building the same piece 4 times with 4 different orientations.\newline
One advantage of this algorithm is that the outer edge for loop can be easily parallelized to achieve better performances.

\section{Testing Results}

In this section the solver that has been proposed by this paper will be tested.
This test will serve two purposes:
Understanding the performance and limitations of the proposed algorithm
Been able to test the same algorithm with both digital, and real world jigsaw puzzles, to understand the differences between the two.

\subsection{Setup}
The test sample consist in 5 different puzzle:
\begin{itemize}
  \item a 4x4 real world puzzle
  \item a 8x8 real world puzzle
  \item a 4x4 digital puzzle
  \item a 8x8 digital puzzle
  \item a 16x16 digital puzzle
\end{itemize}
The real world puzzles have been obtained with a printer’s scanner,
and the digital puzzles have been obtained from \url{https://puzzle.telegnom.org/}.\newline
The constants MIN\_SHORE\_SINGLE\_SIDE and MIN\_SHORE\_PIECE\_GROUP mentioned in section
\ref{document:combining_four_objects} were set at different values
depending on the nature of the puzzle (real or digital).\newline
The test has been performed on a ryzen 5 5600 CPU with 16 GB of RAM.
The time measured in the test does not include the execution time of the Comparator
\ref{document:my_comparator}, but only the one of the Solver~\ref{document:my_solver}.

\subsection{Results}

\begin{table}[H]
  \centering
  \begin{tabular}{
  >{\columncolor[HTML]{D9EAD3}}c 
  >{\columncolor[HTML]{D0E0E3}}c 
  >{\columncolor[HTML]{E6B8AF}}c 
  >{\columncolor[HTML]{FCE5CD}}c }
  \cellcolor[HTML]{B6D7A8} &
    \cellcolor[HTML]{A2C4C9} &
    \multicolumn{2}{c}{\cellcolor[HTML]{EA9999}Execution time {[}s{]}} \\
  \multirow{-2}{*}{\cellcolor[HTML]{B6D7A8}Size} &
    \multirow{-2}{*}{\cellcolor[HTML]{A2C4C9}Pieces} &
    \cellcolor[HTML]{DD7E6B}Digital puzzle &
    \cellcolor[HTML]{F9CB9C}Real Puzzle \\
  4x4   & 16  & 4   & 59   \\
  8x8   & 64  & 30  & 2851 \\
  16x16 & 256 & 377 & N/A 
  \end{tabular}
\end{table}

\subsection{Accuracy}
It is important to note that the project does not have a way to evaluate the accuracy
of the results automatically. But the accuracy for the two real world puzzles
has been asserted manually. And both of them are 100\% correct.

\subsection{Analysis}

This paper has claimed from the start that real world puzzles were harder than digital ones.
Some intuitive explanations were given to justify this claim,
but a formal proof has never been given.
This test finally proves the claim, by showing that the same algorithm can have vastly
different time performances depending on the accuracy of the input data.

It is fair to say that the performance of the algorithm are not amazing,
but they are good enough to achieve the objective of this paper \ref{document:objective}.
It is also worth remembering that the problem itself is extremely complex.
Using the formula introduced in section \ref{document:proof_definitions} we can calculate that a 64 piece puzzle,
like the one that has been solved before, has approximately \(4 \cdot 10^{127}\) possible combinations.
\[C = N! \cdot 4^N =  64! \cdot 4^{64} \approx 4.32 \cdot 10^{127}\]

\section{The Labeling tool}
When the program solves a puzzle it also exports some files that could be used in the future to train
a machine-learning based comparator.
This section is here to explain the data available and the format used.

One piece is uniquely identified by his id piece\_id.
The id is an integer that goes from 0 to Number of pieces - 1.
Each piece has 4 sides that can be identified by the side side\_id,
an integer that goes from 0 to 3.

\subsection{Image of each piece}

Inside the folder results/pieces the program will generate one image
for each one of the pieces of the puzzle. The pieces will be named \{piece\_id\}.jpeg.
An example can be seen in \ref{fig:result_pieces}

\begin{figure}[H]
  \caption{}\label{fig:result_pieces}
  \centering
  \includegraphics[height=0.3\textwidth]{pictures/result_pieces.jpeg}
\end{figure}

\subsection{Image of each side}

Inside the folder results/sides the program will generate one image 
for each side of each piece. The images will be named \{piece\_id\}\_\{side\_id\}.jpeg.
An example can be seen in \ref{fig:result_sides}

\begin{figure}[H]
  \caption{}\label{fig:result_sides}
  \centering
  \includegraphics[height=0.3\textwidth]{pictures/result_sides.jpeg}
\end{figure}

\subsection{Corners of each piece}
The file corners.json will be generated inside the folder results.
This file will contain a list of pieces.
Each piece has 4 sides, and each side has the coordinates of the two points
that delimit itself.\newline
An example of the generated json can be seen below.

\begin{minipage}{\textwidth}
  \begin{lstlisting}
    [
      {
        "piece_id": 0,
        "side_0": {
          "p1": {"x":955, "y":1924},
          "p2": {"x":1024, "y":1016}
        },
        "side_1": {
          "p1": {"x":1024, "y":1016},
          "p2": {"x":1850, "y":1028}
        },
        "side_2": {
          "p1": {"x":1850, "y":1028},
          "p2": {"x":1814, "y":1992}
        },
        "side_3": {
          "p1": {"x":1814, "y":1992},
          "p2": {"x":955, "y":1924}
        }
      },
      {...},
      {...},
      ...
    ]
  \end{lstlisting}
\end{minipage}

\subsection{Match for each side}
The file connections\_result.json will be generated
inside the folder results. This file will contain a list of pieces,
each piece has 4 sides, and each side has a link to the side of the piece
that matches it. The link can be null if the side is along the border.\newline
An example of the generated json can be seen below.

\begin{minipage}{\textwidth}
  \begin{lstlisting}
    [
      {
        "piece_id": 12,
        "side_0": {
            "piece": 5,
            "side": 0
        },
        "side_1": {
            "piece": 8,
            "side": 0
        },
        "side_2": {
            "piece": 0,
            "side": 0
        },
        "side_3": null
      },
      {...},
      {...},
      ...
    ]

  \end{lstlisting}
\end{minipage}

\section{Conclusions}
In conclusion, this paper's objective can be considered achieved.
It has been documented why real-world puzzles are harder than digital puzzles.
It has also been shown that the same algorithm can have drastically worse
performance when working with real-world puzzles compared to digital ones.

A highly effective labeling tool has also been developed.
It could be used in the future to create a machine learning-based comparator that could significantly
improve the time needed to solve this problem by providing more accurate data.

Even the solver that has been created can be considered a success.
His time complexity is horrible if compared with current state-of-the-art digital puzzle solvers.
But if compared with real world one the algorithm is probably one of the best, if not the best, available online.

% referneces page
\clearpage
\begin{thebibliography}{9}

  %Graph Connection Laplacian
  \bibitem{GCL}
    Vahan Huroyan, Gilad   Lerman and Hau-Tieng Wu,
    Solving Jigsaw Puzzles By The Graph Connection Laplacian,
    2020.
    \url{https://arxiv.org/pdf/1811.03188.pdf}.
  % Genetic algorighm
  \bibitem{GA}
    Dror Sholomon, Omid David and Nathan S. Netanyahu,
    A Genetic Algorithm-Based Solver for Very Large Jigsaw Puzzles,
    2013.
    \url{https://openaccess.thecvf.com/content_cvpr_2013/papers/Sholomon_A_Genetic_Algorithm-Based_2013_CVPR_paper.pdf}.

    % Claim of best solution O(N^2)
  \bibitem{ON2Claim}
    Michael Brand,
    No easy puzzles: Hardness results for jigsaw puzzles,
    2015.
    \url{https://www.sciencedirect.com/science/article/pii/S0304397515001607}.

    % abtosoftware real world solution
  \bibitem{Abto}
    AbtoSoftware,
    Computer Vision Powers Automatic Jigsaw Puzzle Solver,
    2019.
    \url{https://www.abtosoftware.com/blog/computer-vision-powers-automatic-jigsaw-puzzle-solver}.

\end{thebibliography}

\end{document}